---
title: "Alpaca"
date: "`r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: hide # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    # toc: true
    # toc_float:
    #   collapsed: true
    #   smooth_scroll: true
    number_sections: false
    fig_caption: true
    code_download: true
always_allow_html: true
bibliography: 
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
---

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
.custom-box {
  background-color: #f5f7fa; /* Light grey-blue background */
  border-color: #e1e8ed; /* Light border color */
  color: #2c3e50; /* Dark text color */
  padding: 15px; /* Padding inside the box */
  border-radius: 5px; /* Rounded corners */
  margin-bottom: 20px; /* Spacing below the box */
}
.caption {
  margin: auto;
  text-align: center;
  margin-bottom: 20px; /* Spacing below the box */
}
```

<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>



```{r, echo= TRUE}
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
# Define the function to truncate a number to two decimal places
truncate_to_two <- function(x) {
  floor(x * 100) / 100
}
```

```{r}
# Load required libraries
library(httr)
library(jsonlite)
library(readr)
library(dplyr)
library(zoo) # to use na.locf
library(tidyr) # to use pivot_wider
library(slackr)
library(grateful)
library(here)
library(DT) 

initial_date <- as.Date("2025-04-23")
source(here("keys.R"))
slackr_setup(token = token)

format_market_cap <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) {
      return("NA")
    }
    
    val <- as.numeric(val)
    
    if (val >= 1e12) {
      return(sprintf("%.1fT", val / 1e12))
    } else if (val >= 1e9) {
      return(sprintf("%.1fB", val / 1e9))
    } else if (val >= 1e6) {
      return(sprintf("%.1fM", val / 1e6))
    } else if (val >= 1e3) {
      return(sprintf("%.1fk", val / 1e3))
    } else {
      return(as.character(val))
    }
  })
}
```

Data is downloaded daily from [here](https://www.nasdaq.com/market-activity/stocks/screener). The `lastsale` value reported in the previous link is the open price when the market closes, according to this [website](https://finance.yahoo.com/).

Data collection started on April 23, 2025.

This code runs daily. So it may happen that the same data is retrieved on both Saturday and Sunday (the data from Friday). So the column `date` is ignored here `distinct(across(-date), .keep_all = TRUE)` because it may be different just because the same data was retrieved in different dates.

```
{r}
aux_day <- Sys.Date()
day_of_week <- weekdays(aux_day)

if (day_of_week == "Monday") {
  today <- aux_day - 3
} else if (day_of_week %in% c("Tuesday", "Wednesday", "Thursday", "Friday")) {
  today <- aux_day - 1
} else {
  today <- aux_day  # or handle Saturday/Sunday however you want
}
# Define the API URL
url <- "https://api.nasdaq.com/api/screener/stocks?tableonly=true&limit=25&offset=0&download=true"
# Set headers to mimic a browser
headers <- add_headers(`User-Agent` = "Mozilla/5.0")
# Send GET request
response <- GET(url, headers)
# Check status
if (status_code(response) == 200) {
  # Parse JSON content
  content_json <- content(response, "text", encoding = "UTF-8")
  data_list <- fromJSON(content_json, flatten = TRUE)

  # Extract rows from JSON
  stock_data <- data_list$data$rows
  # Add current date column
  stock_data$date <- today

  # Create filename with today's date
  file_name <- paste0("nasdaq_screener_", today, ".csv")

  # Save to CSV
  write_csv(stock_data, here("data", file_name))
  cat("✅ Data saved to", file_name, "\n")
} else {
  cat("❌ Failed to fetch data. Status code:", status_code(response), "\n")
}

# The following line should replace the next following line
today_stock_data <- read_csv(here("data", file_name))  |>
  filter(nchar(symbol)<5) |> # filter out stocks with more than 4 characters (not tradeable)
  filter(!grepl("\\^", symbol)) |> # remove symbols containing ^
  select(-url, -netchange, -pctchange) |> # remove unnecessary columns
  mutate(lastsale = as.numeric(gsub("[$,]", "", lastsale))) |># remove $ and , from lastsale and make it numeric
  select(symbol, date, lastsale, everything()) # reorder columns

# Remove symbols that are not longer being traded
cummulative_stock_data <- readRDS(here("data/cummulative_stock_data.RDS")) |>
  filter(symbol %in% today_stock_data$symbol)

# The following two instructions deals with new_data being added when it is already there.
# This might happen when the stock market is closed during weekdays

# Keep only rows from today_stock_data that are NOT already in cummulative_stock_data
new_today_stock_data <- anti_join(today_stock_data, 
                                  cummulative_stock_data, 
                                  by = setdiff(names(today_stock_data), "date")) # Identify columns to match (everything except 'date')

clean_cummulative_stock_data <- bind_rows(cummulative_stock_data, new_today_stock_data) |>
  arrange(symbol, date) |>
  select(symbol, date, lastsale, marketCap, country, industry, sector, ipoyear, volume, name)

# Get unique dates from the dataset
unique_dates <- sort(unique(clean_cummulative_stock_data$date))

# Step 1: Expand each symbol to all existing dates
stock_data_expanded <- clean_cummulative_stock_data |>
  group_by(symbol) |>
  complete(date = unique_dates) |>
  ungroup()

# Step 2: Fill backward for multiple columns
columns_to_fill <- c(
  "lastsale", "marketCap", "country", "industry", 
  "sector", "ipoyear", "volume", "name"
)

stock_data_filled <- stock_data_expanded |>
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  mutate(across(all_of(columns_to_fill),
                ~ zoo::na.locf(.x, fromLast = TRUE, na.rm = FALSE))) |> # filling from the last available value backward
  mutate(across(all_of(columns_to_fill),
                ~ zoo::na.locf(.x, fromLast = FALSE, na.rm = FALSE))) |> # filling from the first available value forward
  ungroup() |>
  arrange(symbol, date)

stock_data_to_save <- stock_data_filled |>
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  filter(!any(is.na(lastsale))) |>
  ungroup() |>
  arrange(symbol, date)

saveRDS(stock_data_to_save, here("data/cummulative_stock_data.RDS"))
```


```{r}
stock_data_to_save <- readRDS(here("data/cummulative_stock_data.RDS")) 
# The first filter ensures that we only keep stocks that have a positive growth trend  
# The second filter ensures that the next day price is at most twice and at least 80% of the last price (from day to day it can only fall 20%)
stocks_data_grow <- stock_data_to_save |>
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  filter(first(lastsale) <= last(lastsale)) |> # ensure positive growth trend
  filter({
    next_adj <- lead(lastsale)
    all((next_adj > 0.8 * lastsale) & (next_adj < 2 * lastsale), na.rm = TRUE)
  }) |>
  ungroup()

# Compute the percentage change from the start date 
# Keep only those stocks where the final price growth is at least half of the highest growth (before the end).
stock_data_pct_change <- stocks_data_grow |>
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  mutate(daily_pct_change = (lastsale - lag(lastsale)) / lag(lastsale) * 100,
         accum_pct_change = (lastsale - first(lastsale)) / first(lastsale) * 100) |>
  filter((max(head(accum_pct_change, -1)) / 2) <= last(accum_pct_change)) |> # max{first, ..., last-1} / 2 <= last
  ungroup() 

# Filter symbols that have fallen (anytime) more that what they have grown (at the end)
symbols_with_negative_pct_change <- stock_data_pct_change |>
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  filter(any(accum_pct_change < 0)) |> # this filter is so that the next make sense
  filter(abs(min(accum_pct_change)) > last(accum_pct_change)) |>
  ungroup() |>
  distinct(symbol) |>
  pull(symbol)

# Filter out those symbols from stock_data_pct_change
# This ensures that have grown more that they have fallen
stock_data_pct_change_that_grow <- stock_data_pct_change |>
  filter(!symbol %in% symbols_with_negative_pct_change)

aux <- stock_data_pct_change_that_grow |>
  filter(date != min(date))


# From here on there are no more filters


# The following contains the information of the surviving symbols
info_of_surviving_symbols <- aux |>
  filter(date == max(date)) |>
  select(-date, -lastsale, -daily_pct_change) |>
  mutate(marketCap_ = format_market_cap(marketCap),
         accum_pct_change = round(accum_pct_change, 2))
  
countries <- info_of_surviving_symbols |>
  select(country) |>
  distinct() |>
  arrange(country) |>
  pull(country)

industries <- info_of_surviving_symbols |>
  select(industry) |>
  distinct() |>
  arrange(industry) |>
  pull(industry)

sectors <- info_of_surviving_symbols |>
  select(sector) |>
  distinct() |>
  arrange(sector) |>
  pull(sector)

# Compute the rank for each date
aux_ranked <- aux |> 
  group_by(date) |>
  arrange(accum_pct_change) |>
  mutate(rank = row_number()) |>
  ungroup() |>
  arrange(symbol,date)

# Compute last day rank
last_day_rank <- aux_ranked |>
  group_by(symbol) |>
  filter(date == max(date)) |>
  summarise(last_rank = rank) |>
  ungroup() |>
  arrange(desc(last_rank))

# Compute the sum of ranks 
sum_rank <- aux_ranked |>
  group_by(symbol) |>
  summarise(sum_rank = sum(rank)) |>
  ungroup() |>
  arrange(desc(sum_rank))

# Create the table of ranks
table_of_ranks <- aux_ranked |>
  select(symbol, date, rank) |>
  pivot_wider(
    names_from = date,
    values_from = rank
  )

# assume first column is 'symbol'
ranking_accumulated <- table_of_ranks

# Get only the numeric part (the ranks)
rank_matrix <- as.matrix(table_of_ranks[,-1])

# Apply the custom accumulation logic row-wise
accumulated_matrix <- t(apply(rank_matrix, 1, function(row) {
  acc <- numeric(length(row))
  acc[1] <- row[1]
  for (i in 2:length(row)) {
    if (row[i] >= row[i - 1]) {
      acc[i] <- acc[i - 1] + row[i]
    } else {
      acc[i] <- acc[i - 1] - row[i]
    }
  }
  return(acc)
}))

# Combine with symbols again
ranking_accumulated[,-1] <- accumulated_matrix

acc_rank <- ranking_accumulated[, c(1,ncol(ranking_accumulated))] |> 
  rename(acc_rank = colnames(ranking_accumulated)[ncol(ranking_accumulated)]) |>
  arrange(desc(acc_rank))


# Augment the ranks with info
aug_acc_rank <- acc_rank |>
  inner_join(info_of_surviving_symbols, by = "symbol") |>
  arrange(desc(acc_rank)) |>
  select(symbol, acc_rank, marketCap, marketCap_, accum_pct_change, country, industry, sector, ipoyear, volume, name) |>
  mutate(symbol = paste0("<a href='https://finance.yahoo.com/quote/", symbol, "' target='_blank'>", symbol, "</a>")) |>
  mutate(rank = row_number()) |>
  select(symbol, acc_rank, marketCap, marketCap_, accum_pct_change, country, industry, sector, ipoyear, volume, rank, name)


aug_sum_rank <- sum_rank |>
  inner_join(info_of_surviving_symbols, by = "symbol") |>
  arrange(desc(sum_rank)) |>
  select(symbol, sum_rank, marketCap, marketCap_, accum_pct_change, country, industry, sector, ipoyear, volume, name) |>
  mutate(symbol = paste0("<a href='https://finance.yahoo.com/quote/", symbol, "' target='_blank'>", symbol, "</a>")) |>
  mutate(rank = row_number()) |>
  select(symbol, sum_rank, marketCap, marketCap_, accum_pct_change, country, industry, sector, ipoyear, volume, rank, name)

aug_last_day_rank <- last_day_rank |>
  inner_join(info_of_surviving_symbols, by = "symbol") |>
  arrange(desc(last_rank)) |>
  select(symbol, last_rank, marketCap, marketCap_, accum_pct_change, country, industry, sector, ipoyear, volume, name) |>
  mutate(symbol = paste0("<a href='https://finance.yahoo.com/quote/", symbol, "' target='_blank'>", symbol, "</a>")) |>
  mutate(rank = row_number()) |>
  select(symbol, last_rank, marketCap, marketCap_, accum_pct_change, country, industry, sector, ipoyear, volume, rank, name)
```

***************

# Accumulated Rank

***************



```{r results='asis'}
datatable(
  aug_acc_rank,
  filter = 'none',
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    dom = 'lrtip',
    initComplete = JS(
  "function(settings) {",
  "  var api = this.api();",
  "  var header = $(api.table().header());",
  "  var filterRow = $('<tr>').appendTo(header);",

  "  var colIndices = { country: 5, industry: 6, sector: 7, ipoyear: 8 };",
  "  var industryToSector = {};",

  "  // First pass: build industry-to-sector map",
  "  var allData = api.rows().data();",
  "  allData.each(function(row) {",
  "    var industry = row[colIndices.industry];",
  "    var sector = row[colIndices.sector];",
  "    if (industry && sector) {",
  "      industryToSector[industry] = sector;",
  "    }",
  "  });",

  "  api.columns().every(function(index) {",
  "    var column = this;",
  "    var cell = $('<th>').appendTo(filterRow);",

  "    if ([colIndices.country, colIndices.industry, colIndices.sector, colIndices.ipoyear].includes(index)) {",
  "      var select = $('<select><option value=\"\"></option></select>')",
  "        .appendTo(cell)",
  "        .on('change', function() {",
  "          var val = $.fn.dataTable.util.escapeRegex($(this).val());",
  "          column.search(val ? '^' + val + '$' : '', true, false).draw();",

  "          if (index === colIndices.sector) {",
  "            var selectedSector = $(this).val();",
  "            var industrySelect = filterRow.find('th').eq(colIndices.industry).find('select');",
  "            industrySelect.val(''); // clear selection",
  "            industrySelect.empty().append('<option value=\"\"></option>');",
  "            var industries = Object.keys(industryToSector).filter(function(ind) {",
  "              return selectedSector === '' || industryToSector[ind] === selectedSector;",
  "            }).sort();",
  "            industries.forEach(function(ind) {",
  "              industrySelect.append('<option value=\"' + ind + '\">' + ind + '</option>');",
  "            });",
  "            // Also clear industry filter when sector is reset",
  "            var industryColumn = api.column(colIndices.industry);",
  "            industryColumn.search('', true, false).draw();",
  "          }",
  "        });",

  "      column.data().unique().sort().each(function(d) {",
  "        select.append('<option value=\"' + d + '\">' + d + '</option>');",
  "      });",
  "    } else if (index === 0) {",
  "      // Add text input for symbol search",
  "      $('<input type=\"text\" placeholder=\"Search symbol\" style=\"width: 100%;\">')",
  "        .appendTo(cell)",
  "        .on('keyup change', function() {",
  "          var val = $(this).val();",
  "          column.search(val).draw();",
  "        });",
  "    } else {",
  "      $(cell).html('');",
  "    }",
  "  });",
  "}"
)
  ),
  escape = FALSE,
  rownames = FALSE,
  class = "display nowrap"
)


```



***************

# Sum of Ranks

***************

```{r}
datatable(
  aug_sum_rank,
  filter = 'none',
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    dom = 'lrtip',
    initComplete = JS(
  "function(settings) {",
  "  var api = this.api();",
  "  var header = $(api.table().header());",
  "  var filterRow = $('<tr>').appendTo(header);",

  "  var colIndices = { country: 5, industry: 6, sector: 7, ipoyear: 8 };",
  "  var industryToSector = {};",

  "  // First pass: build industry-to-sector map",
  "  var allData = api.rows().data();",
  "  allData.each(function(row) {",
  "    var industry = row[colIndices.industry];",
  "    var sector = row[colIndices.sector];",
  "    if (industry && sector) {",
  "      industryToSector[industry] = sector;",
  "    }",
  "  });",

  "  api.columns().every(function(index) {",
  "    var column = this;",
  "    var cell = $('<th>').appendTo(filterRow);",

  "    if ([colIndices.country, colIndices.industry, colIndices.sector, colIndices.ipoyear].includes(index)) {",
  "      var select = $('<select><option value=\"\"></option></select>')",
  "        .appendTo(cell)",
  "        .on('change', function() {",
  "          var val = $.fn.dataTable.util.escapeRegex($(this).val());",
  "          column.search(val ? '^' + val + '$' : '', true, false).draw();",

  "          if (index === colIndices.sector) {",
  "            var selectedSector = $(this).val();",
  "            var industrySelect = filterRow.find('th').eq(colIndices.industry).find('select');",
  "            industrySelect.val(''); // clear selection",
  "            industrySelect.empty().append('<option value=\"\"></option>');",
  "            var industries = Object.keys(industryToSector).filter(function(ind) {",
  "              return selectedSector === '' || industryToSector[ind] === selectedSector;",
  "            }).sort();",
  "            industries.forEach(function(ind) {",
  "              industrySelect.append('<option value=\"' + ind + '\">' + ind + '</option>');",
  "            });",
  "            // Also clear industry filter when sector is reset",
  "            var industryColumn = api.column(colIndices.industry);",
  "            industryColumn.search('', true, false).draw();",
  "          }",
  "        });",

  "      column.data().unique().sort().each(function(d) {",
  "        select.append('<option value=\"' + d + '\">' + d + '</option>');",
  "      });",
  "    } else if (index === 0) {",
  "      // Add text input for symbol search",
  "      $('<input type=\"text\" placeholder=\"Search symbol\" style=\"width: 100%;\">')",
  "        .appendTo(cell)",
  "        .on('keyup change', function() {",
  "          var val = $(this).val();",
  "          column.search(val).draw();",
  "        });",
  "    } else {",
  "      $(cell).html('');",
  "    }",
  "  });",
  "}"
)
  ),
  escape = FALSE,
  rownames = FALSE,
  class = "display nowrap"
)


```

***************

# Last Day Rank

***************

```{r}
datatable(
  aug_last_day_rank,
  filter = 'none',
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    dom = 'lrtip',
    initComplete = JS(
  "function(settings) {",
  "  var api = this.api();",
  "  var header = $(api.table().header());",
  "  var filterRow = $('<tr>').appendTo(header);",

  "  var colIndices = { country: 5, industry: 6, sector: 7, ipoyear: 8 };",
  "  var industryToSector = {};",

  "  // First pass: build industry-to-sector map",
  "  var allData = api.rows().data();",
  "  allData.each(function(row) {",
  "    var industry = row[colIndices.industry];",
  "    var sector = row[colIndices.sector];",
  "    if (industry && sector) {",
  "      industryToSector[industry] = sector;",
  "    }",
  "  });",

  "  api.columns().every(function(index) {",
  "    var column = this;",
  "    var cell = $('<th>').appendTo(filterRow);",

  "    if ([colIndices.country, colIndices.industry, colIndices.sector, colIndices.ipoyear].includes(index)) {",
  "      var select = $('<select><option value=\"\"></option></select>')",
  "        .appendTo(cell)",
  "        .on('change', function() {",
  "          var val = $.fn.dataTable.util.escapeRegex($(this).val());",
  "          column.search(val ? '^' + val + '$' : '', true, false).draw();",

  "          if (index === colIndices.sector) {",
  "            var selectedSector = $(this).val();",
  "            var industrySelect = filterRow.find('th').eq(colIndices.industry).find('select');",
  "            industrySelect.val(''); // clear selection",
  "            industrySelect.empty().append('<option value=\"\"></option>');",
  "            var industries = Object.keys(industryToSector).filter(function(ind) {",
  "              return selectedSector === '' || industryToSector[ind] === selectedSector;",
  "            }).sort();",
  "            industries.forEach(function(ind) {",
  "              industrySelect.append('<option value=\"' + ind + '\">' + ind + '</option>');",
  "            });",
  "            // Also clear industry filter when sector is reset",
  "            var industryColumn = api.column(colIndices.industry);",
  "            industryColumn.search('', true, false).draw();",
  "          }",
  "        });",

  "      column.data().unique().sort().each(function(d) {",
  "        select.append('<option value=\"' + d + '\">' + d + '</option>');",
  "      });",
  "    } else if (index === 0) {",
  "      // Add text input for symbol search",
  "      $('<input type=\"text\" placeholder=\"Search symbol\" style=\"width: 100%;\">')",
  "        .appendTo(cell)",
  "        .on('keyup change', function() {",
  "          var val = $(this).val();",
  "          column.search(val).draw();",
  "        });",
  "    } else {",
  "      $(cell).html('');",
  "    }",
  "  });",
  "}"
)
  ),
  escape = FALSE,
  rownames = FALSE,
  class = "display nowrap"
)

```

```{r}
print(paste("This was executed on:", Sys.time()))
slackr_msg(
  text = "New update at <https://leninrafaelrierasegura.github.io/Alpaca/info.html>",
  channel = "#ranks"
)
```


```{r}
cite_packages(output = "paragraph", out.dir = ".")
```
