---
title: "Alpaca"
date: "`r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: hide # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    # toc: true
    # toc_float:
    #   collapsed: true
    #   smooth_scroll: true
    number_sections: false
    fig_caption: true
    code_download: true
always_allow_html: true
bibliography: 
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
---

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
.custom-box {
  background-color: #f5f7fa; /* Light grey-blue background */
  border-color: #e1e8ed; /* Light border color */
  color: #2c3e50; /* Dark text color */
  padding: 15px; /* Padding inside the box */
  border-radius: 5px; /* Rounded corners */
  margin-bottom: 20px; /* Spacing below the box */
}
.caption {
  margin: auto;
  text-align: center;
  margin-bottom: 20px; /* Spacing below the box */
}
```

<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>



```{r, echo= TRUE}
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
# Define the function to truncate a number to two decimal places
truncate_to_two <- function(x) {
  floor(x * 100) / 100
}
```

```{r}
# Load required libraries
library(httr)
library(jsonlite)
library(readr)
library(dplyr)
library(zoo) # to use na.locf
library(tidyr) # to use pivot_wider
library(slackr)
library(grateful)
library(here)
library(DT) 

initial_date <- as.Date("2025-04-23")
source(here("keys.R"))
slackr_setup(token = token)

format_market_cap <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) {
      return("NA")
    }
    
    val <- as.numeric(val)
    
    if (val >= 1e12) {
      return(sprintf("%.1fT", val / 1e12))
    } else if (val >= 1e9) {
      return(sprintf("%.1fB", val / 1e9))
    } else if (val >= 1e6) {
      return(sprintf("%.1fM", val / 1e6))
    } else if (val >= 1e3) {
      return(sprintf("%.1fk", val / 1e3))
    } else {
      return(as.character(val))
    }
  })
}
```


# Data description


Data is downloaded daily from [here](https://www.nasdaq.com/market-activity/stocks/screener). The `lastsale` value reported in the previous link is the open price when the market closes, according to this [website](https://finance.yahoo.com/).

We only consider data from August 07, 2025. This is the day when the reciprocal US tariffs were put in place.


# Filters

Let `X_k(t_i)` denote the value of a stock `X_k` at day `t_i`, where `i= 0,...,N`. Let `pct X_k(t_i)` denote the percentage change of `X_k(t_i)` with respect to `X_k(t_0)`, i.e., `pct X_k(t_i) = (X_k(t_i) - X_k(t_0))/X_k(t_0) * 100 `. For each stock `X_k`, the filters are as follows:

- The first filter is that `X_k(t_0) <= X_k(t_N)`, i.e., the stock must have a positive growth trend within the period of analysis.

# More stuff

Suppose a stock has price $P_{\mathrm{start}}$ at time $t_{\mathrm{start}}$ and
price $P_{\mathrm{end}}$ at time $t_{\mathrm{end}}$, with
$\Delta t=t_{\mathrm{end}}-t_{\mathrm{start}}$ measured in days. There are
several natural ways to quantify the speed at which the stock has grown.

The simplest measure is the *absolute slope*,
\[
m_{\mathrm{abs}}
=
\frac{P_{\mathrm{end}}-P_{\mathrm{start}}}{\Delta t},
\]
which measures the average increase in price per day. While intuitive,
this measure depends strongly on the stock's initial price, so expensive
stocks tend to have larger values even if their relative growth is modest.

To remove this dependence, one can normalize by the initial price and use
the *fractional slope*,
\[
m_{\mathrm{frac}}
=
\frac{P_{\mathrm{end}}-P_{\mathrm{start}}}
{P_{\mathrm{start}}\Delta t},
\]
which measures the average fractional (or percentage) increase per day.
This metric allows a fair comparison between stocks with different price
levels.

A more natural measure from a financial perspective is the
*log-growth rate*,
\[
m_{\log}
=
\frac{\log(P_{\mathrm{end}}/P_{\mathrm{start}})}
{\Delta t},
\]
which assumes prices grow multiplicatively rather than linearly. If the
stock followed an exponential growth model,
\[
P(t)=P_{\mathrm{start}}e^{m_{\log}t},
\]
then $m_{\log}$ would be the exact continuous growth rate. Moreover,
ranking stocks by $m_{\log}$ is equivalent to ranking them by their
compound annual growth rate (CAGR), making it a widely used measure in
quantitative finance.

As an illustration, consider two stocks observed over the same period of
$100$ days.

As an illustration, consider two stocks observed over the same period of 100 days.

| Stock | Start | End | Days | Absolute slope $\frac{P_{\mathrm{end}}-P_{\mathrm{start}}}{\Delta t}$ | Fractional slope $\frac{P_{\mathrm{end}}-P_{\mathrm{start}}}{P_{\mathrm{start}}\Delta t}$ | Log-growth rate $\frac{\log(P_{\mathrm{end}}/P_{\mathrm{start}})}{\Delta t}$ |
|:------|------:|----:|-----:|----------------:|----------------:|----------------:|
| A | \$1000 | \$1100 | 100 | \$1.00/day | 0.0010/day (0.10%/day) | 0.000953/day |
| B | \$10 | \$30 | 100 | \$0.20/day | 0.0200/day (2.00%/day) | 0.010986/day |

The resulting rankings are

| Metric | Ranking |
|:-------|:--------|
| Absolute slope | $A > B$ |
| Fractional slope | $B > A$ |
| Log-growth rate | $B > A$ |

Although Stock A gains more dollars per day, Stock B grows much faster
relative to its initial value. Consequently, the fractional slope and
log-growth rate provide more meaningful measures of growth when comparing
stocks with different initial prices.


```{r}
stock_data_to_save <- readRDS(here("data/cummulative_stock_data.RDS")) |>
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  filter(date >= as.Date("2026-04-01")) |>
  ungroup() |>
  arrange(symbol, date)

stocks_data_with_comment_column <- stock_data_to_save |> 
  mutate(comment = NA_character_)

# The following ensures that we only keep stocks that have a positive growth trend
stocks_data_grow <- stocks_data_with_comment_column |> 
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  mutate(comment = if_else(
    is.na(comment) & !(first(lastsale) <= last(lastsale)), 
    "did not pass X(t_0) <= X(t_N)", 
    comment
  )) |>
  ungroup()

```

```{r}
slope_comp <- stocks_data_grow |>
  filter(is.na(comment)) |>
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  summarise(
    last_date  = last(date),
    last_price = last(lastsale),
    first_date = first(date),
    first_price = first(lastsale),

    half_date = {
      threshold <- last(lastsale) / 2
      idx <- which(lastsale <= threshold)
      if (length(idx) == 0) as.Date(NA) else max(date[idx])
    },

    days_to_double = as.numeric(last_date - half_date),
    days_from_start = as.numeric(last_date - first_date),

    country  = last(country),
    sector   = last(sector),
    industry = last(industry),
    name     = last(name),
    marketCap = format_market_cap(last(marketCap)),
    ipoyear   = last(ipoyear),

    .groups = "drop"
  ) |>
  mutate(half_price = last_price / 2) |>
  filter(!is.na(half_date)) |>
  arrange(days_to_double) |> 
  mutate(frac_slope = (last_price - first_price) / (first_price * days_from_start)) |>
  mutate(log_slope = log(last_price / first_price) / days_from_start) |>
  select(-last_date, -first_date, -half_date, -half_price, -days_from_start)|>
  relocate(marketCap, frac_slope, log_slope, ipoyear, name, .after = days_to_double)
```

```{r results='asis'}
datatable(
  slope_comp |> mutate(symbol = paste0("<a href='https://finance.yahoo.com/quote/", symbol, "' target='_blank'>", symbol, "</a>")),
  filter = "none",
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    dom = "lrtip",
    initComplete = JS(
      "function(settings) {",
      "  var api = this.api();",
      "  var header = $(api.table().header());",
      "  var filterRow = $('<tr>').appendTo(header);",

      "  var colIndices = { country: 9, sector: 10, industry: 11 };",

      "  var industryToSector = {};",

      "  var allData = api.rows().data();",
      "  allData.each(function(row) {",
      "    var industry = row[colIndices.industry];",
      "    var sector = row[colIndices.sector];",
      "    if (industry && sector) {",
      "      industryToSector[industry] = sector;",
      "    }",
      "  });",

      "  api.columns().every(function(index) {",
      "    var column = this;",
      "    var cell = $('<th>').appendTo(filterRow);",

      "    if ([colIndices.country, colIndices.sector, colIndices.industry].includes(index)) {",

      "      var select = $('<select><option value=\"\"></option></select>')",
      "        .appendTo(cell)",
      "        .on('change', function() {",
      "          var val = $.fn.dataTable.util.escapeRegex($(this).val());",
      "          column.search(val ? '^' + val + '$' : '', true, false).draw();",

      "          if (index === colIndices.sector) {",
      "            var selectedSector = $(this).val();",
      "            var industrySelect = filterRow.find('th').eq(colIndices.industry).find('select');",
      "            industrySelect.val('');",
      "            industrySelect.empty().append('<option value=\"\"></option>');",

      "            var industries = Object.keys(industryToSector)",
      "              .filter(function(ind) {",
      "                return selectedSector === '' || industryToSector[ind] === selectedSector;",
      "              })",
      "              .sort();",

      "            industries.forEach(function(ind) {",
      "              industrySelect.append('<option value=\"' + ind + '\">' + ind + '</option>');",
      "            });",

      "            api.column(colIndices.industry).search('', true, false).draw();",
      "          }",
      "        });",

      "      column.data().unique().sort().each(function(d) {",
      "        select.append('<option value=\"' + d + '\">' + d + '</option>');",
      "      });",

      "    } else if (index === 0) {",

      "      $('<input type=\"text\" placeholder=\"Search symbol\" style=\"width:100%;\">')",
      "        .appendTo(cell)",
      "        .on('keyup change', function() {",
      "          column.search($(this).val()).draw();",
      "        });",

      "    } else {",
      "      $(cell).html('');",
      "    }",
      "  });",
      "}"
    )
  ),
  escape = FALSE,
  rownames = FALSE,
  class = "display nowrap"
)
```

# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```

