---
title: "Alpaca"
date: "`r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: hide # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    # toc: true
    # toc_float:
    #   collapsed: true
    #   smooth_scroll: true
    number_sections: false
    fig_caption: true
    code_download: true
always_allow_html: true
bibliography: 
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
---

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
.custom-box {
  background-color: #f5f7fa; /* Light grey-blue background */
  border-color: #e1e8ed; /* Light border color */
  color: #2c3e50; /* Dark text color */
  padding: 15px; /* Padding inside the box */
  border-radius: 5px; /* Rounded corners */
  margin-bottom: 20px; /* Spacing below the box */
}
.caption {
  margin: auto;
  text-align: center;
  margin-bottom: 20px; /* Spacing below the box */
}
```

<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>



```{r, echo= TRUE}
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
# Define the function to truncate a number to two decimal places
truncate_to_two <- function(x) {
  floor(x * 100) / 100
}
```

```{r}
# Load required libraries
library(httr)
library(jsonlite)
library(readr)
library(dplyr)
library(zoo) # to use na.locf
library(tidyr) # to use pivot_wider
library(slackr)
library(grateful)
library(here)
library(DT) 

initial_date <- as.Date("2025-04-23")
source(here("keys.R"))
slackr_setup(token = token)

format_market_cap <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) {
      return("NA")
    }
    
    val <- as.numeric(val)
    
    if (val >= 1e12) {
      return(sprintf("%.1fT", val / 1e12))
    } else if (val >= 1e9) {
      return(sprintf("%.1fB", val / 1e9))
    } else if (val >= 1e6) {
      return(sprintf("%.1fM", val / 1e6))
    } else if (val >= 1e3) {
      return(sprintf("%.1fk", val / 1e3))
    } else {
      return(as.character(val))
    }
  })
}
```


# Data description


Data is downloaded daily from [here](https://www.nasdaq.com/market-activity/stocks/screener). The `lastsale` value reported in the previous link is the open price when the market closes, according to this [website](https://finance.yahoo.com/).

We only consider data from August 07, 2025. This is the day when the reciprocal US tariffs were put in place.


# Filters

Let `X_k(t_i)` denote the value of a stock `X_k` at day `t_i`, where `i= 0,...,N`. Let `pct X_k(t_i)` denote the percentage change of `X_k(t_i)` with respect to `X_k(t_0)`, i.e., `pct X_k(t_i) = (X_k(t_i) - X_k(t_0))/X_k(t_0) * 100 `. For each stock `X_k`, the filters are as follows:

- The first filter is that `X_k(t_0) <= X_k(t_N)`, i.e., the stock must have a positive growth trend within the period of analysis.

- The second filter is that `0.8*X_k(t_{i-1}) <= X_k(t_i) <= 2*X_k(t_{i-1})`, i.e., the next day price is at most twice and at least 80% of the last price (from day to day it can only fall 20%).

- The third filter is that `max{ pct X_k(t_0), ..., pct X_k(t_{N-1})} / 2 <= pct X_k(t_N)`, i.e., the final price growth is at least half of the highest growth.

- The fourth filter is that `min{|pct X_k(t_0)|, ..., |pct X_k(t_{N-1})|} <= pct X_k(t_N)`, i.e., the stock must not have fallen more than what it has grown.

```{r}
stock_data_to_save <- readRDS(here("data/cummulative_stock_data.RDS")) |>
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  filter(date >= as.Date("2025-08-07")) |>
  ungroup() |>
  arrange(symbol, date)

stocks_data_with_comment_column <- stock_data_to_save |> 
  mutate(comment = NA_character_)

# The following ensures that we only keep stocks that have a positive growth trend
stocks_data_grow <- stocks_data_with_comment_column |> 
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  mutate(comment = if_else(
    is.na(comment) & !(first(lastsale) <= last(lastsale)), 
    "did not pass X(t_0) <= X(t_N)", 
    comment
  )) |>
  ungroup()

# The following ensures that the next day price is at most twice and at least 80% of the last price (from day to day it can only fall 20%)
stocks_data_window <- stocks_data_grow |>
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  mutate(
    comment = if_else(
      is.na(comment) & !all((lead(lastsale) > 0.8 * lastsale) & (lead(lastsale) < 2 * lastsale), na.rm = TRUE),
      "did not pass 0.8*X(t_{i-1}) <= X(t_i) <= 2*X(t_{i-1})", 
      comment
    )) |>
  ungroup()


# Compute the percentage change from the start date (daily and accumulated)
# Keep only those stocks where the final price growth is at least half of the highest growth (before the end).
stock_data_pct_change <- stocks_data_window |>
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  mutate(daily_pct_change = (lastsale - lag(lastsale)) / lag(lastsale) * 100,
         accum_pct_change = (lastsale - first(lastsale)) / first(lastsale) * 100) |>
  mutate(comment = if_else(
    is.na(comment) & !((max(head(accum_pct_change, -1)) / 2) <= last(accum_pct_change)),
    "did not pass max{pct X(t_0), ..., pct X(t_{N-1})} / 2 <= pct X(t_N)", 
    comment
  )) |>
  ungroup() 

# Filter symbols that have fallen (anytime) more that what they have grown (at the end)
stock_data_pct_change_that_grow <- stock_data_pct_change |>
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  mutate(
    comment = if_else(
      is.na(comment) &
      any(accum_pct_change < 0) & !(abs(min(head(accum_pct_change, -1))) <= last(accum_pct_change)),
      "did not pass min{|pct X(t_0)|, ..., |pct X(t_{N-1})|} <= pct X(t_N)",
      comment
    )
  ) |>
  ungroup()

# Removes the first date (the one that was used to compute the daily percentage change)
aux <- stock_data_pct_change_that_grow |>
  filter(date != min(date))

AUX_to_rank <- aux |>
  filter(is.na(comment))

# From here on there are no more filters
```


# Rankings

Let `R_j` be the rank of the stocks according to their percentage change `pct X_k(t_i)` at day `t_i`. The ranks is computed as follows:

- `Accumulated rank`. More complicated than the others. The idea is to compute the rank of each stock at each date, and then accumulate the ranks according to the following logic:

  - If `R_j(t_i) >= R_j(t_{i-1})`, then `R_j(t_i) = R_j(t_{i-1}) + R_j(t_i)`.
  - If `R_j(t_i) < R_j(t_{i-1})`, then `R_j(t_i) = R_j(t_{i-1}) - R_j(t_i)`.

  This means that if a stock has a better rank than the previous day, its accumulated rank increases by the value of its current rank. If it has a worse rank, its accumulated rank decreases by the value of its current rank.

- `Sum of ranks`. This is simply `R_1 + R_2 + ... + R_N`.

- `Last day rank`. This is simply `R_N`.



```{r}
# The following contains the information of the surviving symbols
info_of_surviving_symbols <- aux |>
  filter(date == max(date)) |> # so we get the last market cap and volume, for example
  select(-date, -daily_pct_change) |>
  mutate(marketCap_ = format_market_cap(marketCap),
         accum_pct_change = round(accum_pct_change, 2))

# The following is the list of symbols that passed all filters
info_of_those_that_passed_all_filters <- info_of_surviving_symbols |>
  filter(is.na(comment)) |>
  select(-comment)

# The following is the list of symbols that did not pass all filters
reason_for_those_that_did_not_pass_all_filters <- info_of_surviving_symbols |>
  filter(!is.na(comment)) |>
  select(symbol, comment)
 
# Compute the rank for each date
aux_ranked <- AUX_to_rank |> 
  group_by(date) |>
  arrange(accum_pct_change) |>
  mutate(rank = row_number()) |>
  ungroup() |>
  arrange(symbol,date)

# Compute last day rank
last_day_rank <- aux_ranked |>
  group_by(symbol) |>
  filter(date == max(date)) |>
  summarise(last_rank = rank) |>
  ungroup() |>
  arrange(desc(last_rank))

# Compute the sum of ranks 
sum_rank <- aux_ranked |>
  group_by(symbol) |>
  summarise(sum_rank = sum(rank)) |>
  ungroup() |>
  arrange(desc(sum_rank))

# Create the table of ranks
table_of_ranks <- aux_ranked |>
  select(symbol, date, rank) |>
  pivot_wider(
    names_from = date,
    values_from = rank
  )

# assume first column is 'symbol'
ranking_accumulated <- table_of_ranks

# Get only the numeric part (the ranks)
rank_matrix <- as.matrix(table_of_ranks[,-1])

# Apply the custom accumulation logic row-wise
accumulated_matrix <- t(apply(rank_matrix, 1, function(row) {
  acc <- numeric(length(row))
  acc[1] <- row[1]
  for (i in 2:length(row)) {
    if (row[i] >= row[i - 1]) {
      acc[i] <- acc[i - 1] + row[i]
    } else {
      acc[i] <- acc[i - 1] - row[i]
    }
  }
  return(acc)
}))

# Combine with symbols again
ranking_accumulated[,-1] <- accumulated_matrix

acc_rank <- ranking_accumulated[, c(1,ncol(ranking_accumulated))] |> 
  rename(acc_rank = colnames(ranking_accumulated)[ncol(ranking_accumulated)]) |>
  arrange(desc(acc_rank))


# Augment the ranks with info
aug_acc_rank <- acc_rank |>
  inner_join(info_of_those_that_passed_all_filters, by = "symbol") |>
  arrange(desc(acc_rank)) |>
  select(symbol, acc_rank, marketCap, marketCap_, accum_pct_change, country, industry, sector, ipoyear, volume, lastsale, name) |>
  mutate(rank = row_number()) |>
  mutate(symbol = paste0("<a href='https://finance.yahoo.com/quote/", symbol, "' target='_blank'>", symbol, "</a>")) |>
  select(symbol, acc_rank, marketCap, marketCap_, accum_pct_change, country, industry, sector, ipoyear, volume, rank, lastsale, name)


aug_sum_rank <- sum_rank |>
  inner_join(info_of_those_that_passed_all_filters, by = "symbol") |>
  arrange(desc(sum_rank)) |>
  select(symbol, sum_rank, marketCap, marketCap_, accum_pct_change, country, industry, sector, ipoyear, volume, lastsale, name) |>
  mutate(rank = row_number()) |>
  mutate(symbol = paste0("<a href='https://finance.yahoo.com/quote/", symbol, "' target='_blank'>", symbol, "</a>")) |>
  select(symbol, sum_rank, marketCap, marketCap_, accum_pct_change, country, industry, sector, ipoyear, volume, rank, lastsale, name)

aug_last_day_rank <- last_day_rank |>
  inner_join(info_of_those_that_passed_all_filters, by = "symbol") |>
  arrange(desc(last_rank)) |>
  select(symbol, last_rank, marketCap, marketCap_, accum_pct_change, country, industry, sector, ipoyear, volume, lastsale, name) |>
  mutate(rank = row_number()) |>
  mutate(symbol = paste0("<a href='https://finance.yahoo.com/quote/", symbol, "' target='_blank'>", symbol, "</a>")) |>
  select(symbol, last_rank, marketCap, marketCap_, accum_pct_change, country, industry, sector, ipoyear, volume, rank, lastsale, name)
```

***************

# Accumulated Rank

***************



```{r results='asis'}
datatable(
  aug_acc_rank,
  filter = 'none',
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    dom = 'lrtip',
    initComplete = JS(
  "function(settings) {",
  "  var api = this.api();",
  "  var header = $(api.table().header());",
  "  var filterRow = $('<tr>').appendTo(header);",

  "  var colIndices = { country: 5, industry: 6, sector: 7, ipoyear: 8 };",
  "  var industryToSector = {};",

  "  // First pass: build industry-to-sector map",
  "  var allData = api.rows().data();",
  "  allData.each(function(row) {",
  "    var industry = row[colIndices.industry];",
  "    var sector = row[colIndices.sector];",
  "    if (industry && sector) {",
  "      industryToSector[industry] = sector;",
  "    }",
  "  });",

  "  api.columns().every(function(index) {",
  "    var column = this;",
  "    var cell = $('<th>').appendTo(filterRow);",

  "    if ([colIndices.country, colIndices.industry, colIndices.sector, colIndices.ipoyear].includes(index)) {",
  "      var select = $('<select><option value=\"\"></option></select>')",
  "        .appendTo(cell)",
  "        .on('change', function() {",
  "          var val = $.fn.dataTable.util.escapeRegex($(this).val());",
  "          column.search(val ? '^' + val + '$' : '', true, false).draw();",

  "          if (index === colIndices.sector) {",
  "            var selectedSector = $(this).val();",
  "            var industrySelect = filterRow.find('th').eq(colIndices.industry).find('select');",
  "            industrySelect.val(''); // clear selection",
  "            industrySelect.empty().append('<option value=\"\"></option>');",
  "            var industries = Object.keys(industryToSector).filter(function(ind) {",
  "              return selectedSector === '' || industryToSector[ind] === selectedSector;",
  "            }).sort();",
  "            industries.forEach(function(ind) {",
  "              industrySelect.append('<option value=\"' + ind + '\">' + ind + '</option>');",
  "            });",
  "            // Also clear industry filter when sector is reset",
  "            var industryColumn = api.column(colIndices.industry);",
  "            industryColumn.search('', true, false).draw();",
  "          }",
  "        });",

  "      column.data().unique().sort().each(function(d) {",
  "        select.append('<option value=\"' + d + '\">' + d + '</option>');",
  "      });",
  "    } else if (index === 0) {",
  "      // Add text input for symbol search",
  "      $('<input type=\"text\" placeholder=\"Search symbol\" style=\"width: 100%;\">')",
  "        .appendTo(cell)",
  "        .on('keyup change', function() {",
  "          var val = $(this).val();",
  "          column.search(val).draw();",
  "        });",
  "    } else {",
  "      $(cell).html('');",
  "    }",
  "  });",
  "}"
)
  ),
  escape = FALSE,
  rownames = FALSE,
  class = "display nowrap"
)


```



***************

# Sum of Ranks

***************

```{r}
datatable(
  aug_sum_rank,
  filter = 'none',
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    dom = 'lrtip',
    initComplete = JS(
  "function(settings) {",
  "  var api = this.api();",
  "  var header = $(api.table().header());",
  "  var filterRow = $('<tr>').appendTo(header);",

  "  var colIndices = { country: 5, industry: 6, sector: 7, ipoyear: 8 };",
  "  var industryToSector = {};",

  "  // First pass: build industry-to-sector map",
  "  var allData = api.rows().data();",
  "  allData.each(function(row) {",
  "    var industry = row[colIndices.industry];",
  "    var sector = row[colIndices.sector];",
  "    if (industry && sector) {",
  "      industryToSector[industry] = sector;",
  "    }",
  "  });",

  "  api.columns().every(function(index) {",
  "    var column = this;",
  "    var cell = $('<th>').appendTo(filterRow);",

  "    if ([colIndices.country, colIndices.industry, colIndices.sector, colIndices.ipoyear].includes(index)) {",
  "      var select = $('<select><option value=\"\"></option></select>')",
  "        .appendTo(cell)",
  "        .on('change', function() {",
  "          var val = $.fn.dataTable.util.escapeRegex($(this).val());",
  "          column.search(val ? '^' + val + '$' : '', true, false).draw();",

  "          if (index === colIndices.sector) {",
  "            var selectedSector = $(this).val();",
  "            var industrySelect = filterRow.find('th').eq(colIndices.industry).find('select');",
  "            industrySelect.val(''); // clear selection",
  "            industrySelect.empty().append('<option value=\"\"></option>');",
  "            var industries = Object.keys(industryToSector).filter(function(ind) {",
  "              return selectedSector === '' || industryToSector[ind] === selectedSector;",
  "            }).sort();",
  "            industries.forEach(function(ind) {",
  "              industrySelect.append('<option value=\"' + ind + '\">' + ind + '</option>');",
  "            });",
  "            // Also clear industry filter when sector is reset",
  "            var industryColumn = api.column(colIndices.industry);",
  "            industryColumn.search('', true, false).draw();",
  "          }",
  "        });",

  "      column.data().unique().sort().each(function(d) {",
  "        select.append('<option value=\"' + d + '\">' + d + '</option>');",
  "      });",
  "    } else if (index === 0) {",
  "      // Add text input for symbol search",
  "      $('<input type=\"text\" placeholder=\"Search symbol\" style=\"width: 100%;\">')",
  "        .appendTo(cell)",
  "        .on('keyup change', function() {",
  "          var val = $(this).val();",
  "          column.search(val).draw();",
  "        });",
  "    } else {",
  "      $(cell).html('');",
  "    }",
  "  });",
  "}"
)
  ),
  escape = FALSE,
  rownames = FALSE,
  class = "display nowrap"
)


```

***************

# Last Day Rank

***************

```{r}
datatable(
  aug_last_day_rank,
  filter = 'none',
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    dom = 'lrtip',
    initComplete = JS(
  "function(settings) {",
  "  var api = this.api();",
  "  var header = $(api.table().header());",
  "  var filterRow = $('<tr>').appendTo(header);",

  "  var colIndices = { country: 5, industry: 6, sector: 7, ipoyear: 8 };",
  "  var industryToSector = {};",

  "  // First pass: build industry-to-sector map",
  "  var allData = api.rows().data();",
  "  allData.each(function(row) {",
  "    var industry = row[colIndices.industry];",
  "    var sector = row[colIndices.sector];",
  "    if (industry && sector) {",
  "      industryToSector[industry] = sector;",
  "    }",
  "  });",

  "  api.columns().every(function(index) {",
  "    var column = this;",
  "    var cell = $('<th>').appendTo(filterRow);",

  "    if ([colIndices.country, colIndices.industry, colIndices.sector, colIndices.ipoyear].includes(index)) {",
  "      var select = $('<select><option value=\"\"></option></select>')",
  "        .appendTo(cell)",
  "        .on('change', function() {",
  "          var val = $.fn.dataTable.util.escapeRegex($(this).val());",
  "          column.search(val ? '^' + val + '$' : '', true, false).draw();",

  "          if (index === colIndices.sector) {",
  "            var selectedSector = $(this).val();",
  "            var industrySelect = filterRow.find('th').eq(colIndices.industry).find('select');",
  "            industrySelect.val(''); // clear selection",
  "            industrySelect.empty().append('<option value=\"\"></option>');",
  "            var industries = Object.keys(industryToSector).filter(function(ind) {",
  "              return selectedSector === '' || industryToSector[ind] === selectedSector;",
  "            }).sort();",
  "            industries.forEach(function(ind) {",
  "              industrySelect.append('<option value=\"' + ind + '\">' + ind + '</option>');",
  "            });",
  "            // Also clear industry filter when sector is reset",
  "            var industryColumn = api.column(colIndices.industry);",
  "            industryColumn.search('', true, false).draw();",
  "          }",
  "        });",

  "      column.data().unique().sort().each(function(d) {",
  "        select.append('<option value=\"' + d + '\">' + d + '</option>');",
  "      });",
  "    } else if (index === 0) {",
  "      // Add text input for symbol search",
  "      $('<input type=\"text\" placeholder=\"Search symbol\" style=\"width: 100%;\">')",
  "        .appendTo(cell)",
  "        .on('keyup change', function() {",
  "          var val = $(this).val();",
  "          column.search(val).draw();",
  "        });",
  "    } else {",
  "      $(cell).html('');",
  "    }",
  "  });",
  "}"
)
  ),
  escape = FALSE,
  rownames = FALSE,
  class = "display nowrap"
)

```

***************

# Reason for ranking dismissal

***************

```{r}
datatable(
  reason_for_those_that_did_not_pass_all_filters,
  filter = "top",         # Show filter boxes on top of the columns
  options = list(
    pageLength = 5,
    autoWidth = TRUE,
    columnDefs = list(
      list(targets = 1, searchable = TRUE)  # Disable search on the 'comment' column
    )
  )
)
```



```{r}
range_of_analysis <-range(stock_data_to_save$date)
print(paste("This was executed on:", Sys.time(), 
            "and the range of analysis is from", 
            range_of_analysis[1], 
            "to", 
            range_of_analysis[2]))
```



```{r}
cite_packages(output = "paragraph", out.dir = ".")
```
