---
title: "Fastest doubling stocks"
date: "`r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: hide # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    # toc: true
    # toc_float:
    #   collapsed: true
    #   smooth_scroll: true
    number_sections: false
    fig_caption: true
    code_download: true
always_allow_html: true
bibliography: 
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
---

```{r xaringanExtra-clipboard, echo = FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


```{css, echo = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
   font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
.custom-box {
  background-color: #f5f7fa; /* Light grey-blue background */
  border-color: #e1e8ed; /* Light border color */
  color: #2c3e50; /* Dark text color */
  padding: 15px; /* Padding inside the box */
  border-radius: 5px; /* Rounded corners */
  margin-bottom: 20px; /* Spacing below the box */
}
.caption {
  margin: auto;
  text-align: center;
  margin-bottom: 20px; /* Spacing below the box */
}
```

<div style="color: blue;">
********
**Press the Show button below to reveal the code.**

********
</div>



```{r, echo= TRUE}
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
# Define the function to truncate a number to two decimal places
truncate_to_two <- function(x) {
  floor(x * 100) / 100
}
```

```{r}
# Load required libraries
library(httr)
library(jsonlite)
library(readr)
library(dplyr)
library(zoo) # to use na.locf
library(tidyr) # to use pivot_wider
library(slackr)
library(grateful)
library(here)
library(DT) 

initial_date <- as.Date("2025-04-23")
source(here("keys.R"))
slackr_setup(token = token)

format_market_cap <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) {
      return("NA")
    }
    
    val <- as.numeric(val)
    
    if (val >= 1e12) {
      return(sprintf("%.1fT", val / 1e12))
    } else if (val >= 1e9) {
      return(sprintf("%.1fB", val / 1e9))
    } else if (val >= 1e6) {
      return(sprintf("%.1fM", val / 1e6))
    } else if (val >= 1e3) {
      return(sprintf("%.1fk", val / 1e3))
    } else {
      return(as.character(val))
    }
  })
}
```


# Data description


Data is downloaded daily from [here](https://www.nasdaq.com/market-activity/stocks/screener). The `lastsale` value reported in the previous link is the open price when the market closes, according to this [website](https://finance.yahoo.com/).

We only consider data from August 07, 2025. This is the day when the reciprocal US tariffs were put in place.


# Filters

Let `X_k(t_i)` denote the value of a stock `X_k` at day `t_i`, where `i= 0,...,N`. Let `pct X_k(t_i)` denote the percentage change of `X_k(t_i)` with respect to `X_k(t_0)`, i.e., `pct X_k(t_i) = (X_k(t_i) - X_k(t_0))/X_k(t_0) * 100 `. For each stock `X_k`, the filters are as follows:

- The first filter is that `X_k(t_0) <= X_k(t_N)`, i.e., the stock must have a positive growth trend within the period of analysis.


```{r}
stock_data_to_save <- readRDS(here("data/cummulative_stock_data.RDS")) |>
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  filter(date >= as.Date("2025-08-07")) |>
  ungroup() |>
  arrange(symbol, date)

stocks_data_with_comment_column <- stock_data_to_save |> 
  mutate(comment = NA_character_)

# The following ensures that we only keep stocks that have a positive growth trend
stocks_data_grow <- stocks_data_with_comment_column |> 
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  mutate(comment = if_else(
    is.na(comment) & !(first(lastsale) <= last(lastsale)), 
    "did not pass X(t_0) <= X(t_N)", 
    comment
  )) |>
  ungroup()

```

```{r}
doubling_speed <- stocks_data_grow |>
  filter(is.na(comment)) |>
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  summarise(
    last_date  = last(date),
    last_price = last(lastsale),

    half_date = {
      threshold <- last(lastsale) / 2
      idx <- which(lastsale <= threshold)
      if (length(idx) == 0) as.Date(NA) else max(date[idx])
    },

    days_to_double = as.numeric(last_date - half_date),

    country  = last(country),
    sector   = last(sector),
    industry = last(industry),
    name     = last(name),
    marketCap = format_market_cap(last(marketCap)),
    ipoyear   = last(ipoyear),

    .groups = "drop"
  ) |>
  mutate(half_price = last_price / 2) |>
  filter(!is.na(half_date)) |>
  arrange(days_to_double) 
```

```{r results='asis'}
datatable(
  doubling_speed |> mutate(symbol = paste0("<a href='https://finance.yahoo.com/quote/", symbol, "' target='_blank'>", symbol, "</a>")),
  filter = "none",
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    dom = "lrtip",
    initComplete = JS(
      "function(settings) {",
      "  var api = this.api();",
      "  var header = $(api.table().header());",
      "  var filterRow = $('<tr>').appendTo(header);",

      "  var colIndices = { country: 5, sector: 6, industry: 7 };",

      "  var industryToSector = {};",

      "  var allData = api.rows().data();",
      "  allData.each(function(row) {",
      "    var industry = row[colIndices.industry];",
      "    var sector = row[colIndices.sector];",
      "    if (industry && sector) {",
      "      industryToSector[industry] = sector;",
      "    }",
      "  });",

      "  api.columns().every(function(index) {",
      "    var column = this;",
      "    var cell = $('<th>').appendTo(filterRow);",

      "    if ([colIndices.country, colIndices.sector, colIndices.industry].includes(index)) {",

      "      var select = $('<select><option value=\"\"></option></select>')",
      "        .appendTo(cell)",
      "        .on('change', function() {",
      "          var val = $.fn.dataTable.util.escapeRegex($(this).val());",
      "          column.search(val ? '^' + val + '$' : '', true, false).draw();",

      "          if (index === colIndices.sector) {",
      "            var selectedSector = $(this).val();",
      "            var industrySelect = filterRow.find('th').eq(colIndices.industry).find('select');",
      "            industrySelect.val('');",
      "            industrySelect.empty().append('<option value=\"\"></option>');",

      "            var industries = Object.keys(industryToSector)",
      "              .filter(function(ind) {",
      "                return selectedSector === '' || industryToSector[ind] === selectedSector;",
      "              })",
      "              .sort();",

      "            industries.forEach(function(ind) {",
      "              industrySelect.append('<option value=\"' + ind + '\">' + ind + '</option>');",
      "            });",

      "            api.column(colIndices.industry).search('', true, false).draw();",
      "          }",
      "        });",

      "      column.data().unique().sort().each(function(d) {",
      "        select.append('<option value=\"' + d + '\">' + d + '</option>');",
      "      });",

      "    } else if (index === 0) {",

      "      $('<input type=\"text\" placeholder=\"Search symbol\" style=\"width:100%;\">')",
      "        .appendTo(cell)",
      "        .on('keyup change', function() {",
      "          column.search($(this).val()).draw();",
      "        });",

      "    } else {",
      "      $(cell).html('');",
      "    }",
      "  });",
      "}"
    )
  ),
  escape = FALSE,
  rownames = FALSE,
  class = "display nowrap"
)
```

```{r}
library(dplyr)
library(plotly)
library(purrr)

plot_list <- map(
  seq(1, 181, by = 20),
  function(start_rank) {

    end_rank <- start_rank + 19

    ranked_stocks <- doubling_speed |>
      slice(start_rank:end_rank) |>
      select(symbol, days_to_double)

    plot_data <- stocks_data_grow |>
      filter(symbol %in% ranked_stocks$symbol) |>
      left_join(ranked_stocks, by = "symbol") |>
      mutate(
        symbol = factor(
          symbol,
          levels = ranked_stocks$symbol
        )
      )

    plot_ly(
      data = plot_data,
      x = ~date,
      y = ~lastsale,
      color = ~symbol,
      type = "scatter",
      mode = "lines",
      hovertemplate = paste(
        "<b>%{fullData.name}</b><br>",
        "Date: %{x}<br>",
        "Price: %{y:.2f}<br>",
        "Days to double: %{customdata}<extra></extra>"
      ),
      customdata = ~days_to_double
    ) |>
      layout(
        title = paste(
          "Fastest Doubling Stocks Ranked",
          start_rank, "to", end_rank
        ),
        xaxis = list(title = "Date"),
        yaxis = list(title = "Price"),
        legend = list(
          title = list(text = "Symbol")
        )
      )
  }
)
```

## Fastest doubling stocks {.tabset}

### Rank 1-20

```{r, out.width='50%'}
plot_list[[1]]
```


### Rank 21-40

```{r, out.width='50%'}
plot_list[[2]]
```


### Rank 41-60

```{r, out.width='50%'}
plot_list[[3]]
```

### Rank 61-80

```{r, out.width='50%'}
plot_list[[4]]
```


### Rank 81-100

```{r, out.width='50%'}
plot_list[[5]]
```

### Rank 101-120

```{r, out.width='50%'}
plot_list[[6]]
```


### Rank 121-140

```{r, out.width='50%'}
plot_list[[7]]
```

### Rank 141-160

```{r, out.width='50%'}
plot_list[[8]]
```

### Rank 161-180

```{r, out.width='50%'}
plot_list[[9]]
```

### Rank 181-200

```{r, out.width='50%'}
plot_list[[10]]
```

# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```

